Normalization Circularity in Population-Based FA Analysis

### 2011-06-07 ANTsTBSS/paper.tex
ENH: one specific sentence in discussion

\subsection{Discussion} Prior evaluation work using anatomically labeled
data \cite{bunch_of_stuff} indicates that mutual information,
cross-correlation and the SSD/Demons metric are all capable of producing
high quality anatomical alignment.  All of this work, though, compares
the quality of alignment at a relatively coarse scale of major gyri,
lobes and regions.  At the voxel level, it is nearly impossible to
determine the ground truth correspondence.  Indeed, there is no reason
to believe that subtle differences between a set of alignment strategies
(induced, in the experiments of this paper, by changing the similarity
metrics used in image registration) that all apparently work well should
be considered significant.  This work, however, highlights a dramatic impact of similarity metric choice on detection power in template-based FA studies.  Our contention is that this dramatic "improvement" in detection power is not due to better anatomical alignment.  Rather, it is a result of circularity/non-independence of the normalization and statistical estimation strategy.  Consequently, we recommend that future work should take greater care when pairing the normalization strategy with the statistical framework.  Furthermore, we recommend that researchers should diligently seek to maintain independence between these two critical stages of a population study.  That is, {\textcolor{red}{{\em the features that drive the minimization of a similarity metric should be independent of the features that will be used in hypothesis testing.)}}}

>>>>

\subsection{Discussion} Prior evaluation work using anatomically labeled
data \cite{bunch_of_stuff} indicates that mutual information,
cross-correlation and the SSD/Demons metric are all capable of producing
high quality anatomical alignment.  All of this work, though, compares
the quality of alignment at a relatively coarse scale of major gyri,
lobes and regions.  At the voxel level, it is nearly impossible to
determine the ground truth correspondence.  Indeed, there is no reason
to believe that subtle differences between a set of alignment strategies
(induced, in the experiments of this paper, by changing the similarity
metrics used in image registration) that all apparently work well should
be considered significant.  This work, however, highlights a dramatic impact of similarity metric choice on detection power in template-based FA studies.  Our contention is that this dramatic "improvement" in detection power is not due to better anatomical alignment.  Rather, it is a result of circularity/non-independence of the normalization and statistical estimation strategy.  Consequently, we recommend that future work should take greater care when pairing the normalization strategy with the statistical framework.  Furthermore, we recommend that researchers should diligently seek to maintain independence between these two critical stages of a population study.  That is, {\textcolor{red}{{\em the features that drive the minimization of a similarity metric should be independent of the features that will be used in hypothesis testing.)}}}  In particular, one should avoid combining the SSD metric for normalizing FA with Student's t-test for FA differences.
